---
title: "SMPA - Grouping Assignment"
author: "YourNameHere"
date: "`r format(Sys.time(), '%B %d, %Y')`"
format:
  html:
    self-contained: true
---

```{r}

#load libraries
library(tidyverse)
library(lubridate)

#load data
payments <- readRDS("data/trump_campaign_payments.rds")

```

## Introduction

**What data will you be working with?**

A dataset of disbursements (payments and purchases) by the Donald J. Trump for President campaign committee during the 2020 election cycle. The included payments are from 2017 and 2018.

This data and examples were developed as part of the NYT Open training, which we've adapted from here. Note that the dataset has been alterted for instructional purposes, so doesn't represent the original FEC data in its entirety.

In the environment tab in the upper right, click on the `payments` dataset to open the RStudio viewer to examine the data. You can also run the chunk below as well to bring up a slice of the data.

```{r}

head(payments)


```

## Before you begin

Take a moment to look at the various categories included for each disbursement.

-   Notice that there are several date-related fields:

    -   FEC election cycle

    -   report year

    -   election type

    -   disbursement date

-   There are a few fields about the recipient. Along with city and state, you have:

    -   the recipient name

    -   the payee's first and last name (an individual who was paid for the "recipient") and

    -   the entity type (individual v. organization)

## Initial Narrative Questions

Without doing any analysis yet, what are three questions you might try to answer using this dataset? Write your narrative answers in the spaces below.

1.  Which cycle did the most disbursements occur in?

2.  What type of services did the campaign committee pay out the most to?

3.  Did the campaign pay out to the same recipient numerous times?

## Coding Assignment Questions

**Please write your R tidyverse code to help answer the questions below in the provided R chunks. You can make more than one chunk if you prefer.**

**Also, describe what you're doing either in text out "in the open" above the chunk, or in comments (prefaced by #) in the R chunk itself.**

**You'll need to use grouping, filtering, or a combination of both to find the answers. Remember that R is case-sensitive when you refer to any text values within the columns**

<br>

How much was spent in the 2017 vs. 2018 reporting year?
Answer: There was $1688149 spent in 2017 vs. $37406905 spent in 2018.

```{r}

payments %>% 
  group_by(report_year) %>% 
  summarise(total_amt = sum(disbursement_amount))





```

Tally the total overall disbursement amounts for each month (distinguishing between say, Oct. 2018 v. Oct. 2019). In which month was spending the highest?
Answer: The most money spent was in October of 2018 with a total amount of $9738567.

```{r}

payments %>% 
  group_by(disbursement_year, disbursement_month) %>% 
  summarise(total_amt = sum(disbursement_amount)) %>% 
  arrange(desc(total_amt))



```

Which recipient (recipient_name) was paid the most in the 2020 election cycle?
Answer: The recipient with the most money paid out to them during the 2020 election cycle was American Made Media Consultants, LLC.

```{r}

payments %>%
  filter(election_cycle == "2020") %>% 
  group_by(recipient_name) %>% 
  summarise(total_amt = sum(disbursement_amount)) %>% 
  arrange(desc(total_amt))






```

Look at what this recipient was paid for. What service or good (see the "disbursement_description") were they paid the most for?
Answer: The most service that American Made Media Consultants, LLC was paid the most for was "Placed Media".
```{r}
payments %>% 
  filter(recipient_name == "AMERICAN MADE MEDIA CONSULTANTS, LLC") %>% 
  group_by(disbursement_description) %>% 
  summarise(total_amt = sum(disbursement_amount)) %>% 
  arrange(desc(total_amt))
  



```

What's the most common category overall (see "disbursement_description") the campaign spent money on?
Answer: The most common category for payments was Merchant Fees.
```{r}

payments %>% 
  group_by(disbursement_description) %>% 
  summarise(count = n()) %>%
  arrange(desc(count))




```

Who were the 5 biggest recipients of campaign spending on "LEGAL CONSULTING"..?
Answer: The five biggest recipients for legal consulting were Jones Day, Larocca Hornik Rosen Greenberg, Kasowits Benson Torress, Law Offices of Alan S. Futerfas, and McDermott Will & Emery.
```{r}

payments %>% 
  filter(disbursement_description == "LEGAL CONSULTING") %>% 
  group_by(recipient_name) %>% 
  summarise(total_amt = sum(disbursement_amount)) %>% 
  arrange(desc(total_amt))
  





```


Who are the top 5 individuals (vs. organization) were paid by the campaign committee?
Answer: The top five individuals by total amount paid were Michael Glassner, Sean Dollman, Darren Centinello, Christopher Carr, and John Pence.
```{r}

payments %>% 
  filter(entity_type == "IND") %>% 
  group_by(recipient_name) %>% 
  summarise(total_amt = sum(disbursement_amount)) %>% 
  arrange(desc(total_amt))






```


For the indvidual who was paid the most, which payment category (disbursement_description) was the top one for him or her? And how much money were they paid in that category?
Answer: For Michael Glassner, the most common category was payroll. He was paid out for payroll 47 times, amounting to a total of $10,000.
```{r}

payments %>% 
  filter(recipient_name == "GLASSNER, MICHAEL") %>% 
  group_by(disbursement_description, disbursement_amount) %>% 
  summarise(count = n()) %>%
  arrange(desc(count))





```
